---
title: "Figure_2_Tuft_ADM.Rmd"
author: "Hue M. La"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, error = F, message = F, results = "hide", eval = T, fig.align="center")
```

# Installing packages
```{r Packages}
library(Seurat)
library(ggplot2)
library(shiny)
library(dplyr)
library(patchwork)
library(rmarkdown)
library(openxlsx)
library(here)

objects.path <- here("output/objects/")
data.path <- here("data/")
```

# Data loading 
```{r eval=FALSE}
sts.non.malignant <- readRDS("../objects/sts.nonmalignant.RDS")
```

# TUFT CELL ANALYSIS
## Reclustered of Tuft to identify persistent Tuft
```{r}
tuft <- subset(sts.non.malignant, subset = Lv.1.Anno == "Tuft")
table(tuft$Patient_ID)

# Combined samples with small Tuft < 5 cells
# Reassign cluster names 
metadata <- tuft@meta.data %>% as.data.frame() #Creating a data frame called metadata
metadata$S_CB <- rownames(metadata)#Assigning S_CB as a new column (CB=cell barcode)
metadata %>% mutate(Patient_c = case_when(Patient_ID  == "2156" ~ "2156",
                                          Patient_ID  == "2215" ~ "2215",
                                          Patient_ID  == "2221" ~ "2221",
                                          Patient_ID  == "2285" ~ "2285",
                                          Patient_ID  == "2925" ~ "2925",
                                          TRUE ~ "4Pats" )) -> metadata
metadata %>% group_by(Patient_ID) %>% count(Patient_c)
Patient_c <- metadata$Patient_c 
names(Patient_c) <- metadata$S_CB
tuft <- AddMetaData(object = tuft, metadata = Patient_c, col.name = "Patient_c")

# DietSeurat
colnames(tuft[[]]) #check which columns to remove
tuft[["nCount_SCT"]] <- NULL
tuft[["nFeature_SCT"]] <- NULL
tuft[["seurat_clusters"]] <- NULL
tuft[["SCT_snn_res.0.6"]] <- NULL
tuft[["SCT_snn_res.0.7"]] <- NULL
tuft[["SCT_snn_res.0.8"]] <- NULL

# Due to small number of cells < 30 -> can't do integration
tuft <- SplitObject(tuft, split.by = "Patient_c") 

tuft <- lapply(tuft, FUN= SCTransform, vst.flavor="v2", assay = "RNA", method = "glmGamPoi",
              variable.features.n = 2000)

features.split <- SelectIntegrationFeatures(object.list = tuft, nfeatures = 2000)

tuft.merged <- merge(x=tuft[[1]], y=tuft[2:length(tuft)])
VariableFeatures(tuft.merged) <- features.split

# Dimentionality reduction
tuft.merged <- RunPCA(tuft.merged, npcs = 30, verbose = FALSE, assay = "SCT")
ElbowPlot(tuft.merged)

DimPlot(tuft.merged, label = TRUE,
        reduction = "pca", group.by = "Patient_c") + NoLegend()

tuft.merged <- RunUMAP(tuft.merged, reduction = "pca", dims = 1:30)

# Clustering
tuft.merged <- FindNeighbors(tuft.merged, reduction = "pca", dims = 1:30)

# Determine cluster stability
tuft.merged <- FindClusters(tuft.merged, algorithm = 3,  resolution = seq(0.1, 1, by = 0.1))

library(clustree)
clustree::clustree(tuft.merged) 

DimPlot(tuft.merged, group.by = "SCT_snn_res.0.9") + #DimPlot(tuft.merged, group.by = "SCT_snn_res.1") + 
  DimPlot(tuft.merged, group.by = "Patient_c")
saveRDS(tuft.merged, file= here(objects.path, "tuft.merged.RDS"))
```

## Reclustered of Tuft based on 4 ADM samples only
```{r}
tuft.merged <- subset(tuft.merged, subset = Patient_c %in% c("4Pats","2925"), invert = T)
table(tuft$Patient_ID)

# FindMarkers 
Idents(tuft.merged) <- "Lv.2.Anno"
tuft.markers <- FindAllMarkers(tuft.merged, assay = "RNA", min.pct = 0.15, logfc.threshold = 0.25)
markers <- tuft.markers %>% group_by(cluster) %>% slice_max(n=30, order_by = avg_log2FC) %>% pull(gene) %>% unique()
cluster_anno <- tuft.merged@meta.data %>% dplyr::select(Lv.2.Anno, Patient_ID) %>% as.data.frame()
tuft.col <- c("#6A3D9A" ,"#D8AA57")
pat.col <- c("#DE73DA" ,"#E7E3DA" ,"#67C0D8" ,"#E1A497" )  #
cluster_colors <- list(Lv.2.Anno = tuft.col, Patient_ID = pat.col)
names(cluster_colors$Lv.2.Anno) <- paste0(c("Tuft","Tuft.persistent"))
names(cluster_colors$Patient_ID) <- paste0(c( "2156","2215","2221", "2285"))

mat3 <- tuft.merged[["RNA"]]@data[markers, ] %>% as.matrix()
mat3 <- t(scale(t(mat3)))

pheatmap::pheatmap(mat3, color=colorRampPalette(c("steelblue","white","darkred"))(100), border_color="NA",
    breaks=seq(-1.5,1.5,length.out=101), cluster_cols=TRUE, scale="none", fontsize_row=7, 
    show_colnames=FALSE, treeheight_row=70, treeheight_col=70, cutree_cols=2,
    clustering_method="ward.D2", annotation_col=cluster_anno, annotation_colors=cluster_colors)
```

## Confirming tuft mutation status for sample 2221_ICGC_0105
```{r}
tuft.cnv <- read_csv("~/projects/PDAC_non_malignant_cell_types/data/import.files/tuft.snv.df.csv")
names(tuft.cnv)[1] <- "Barcode"
kras.mut <- tuft.cnv %>% dplyr::select("Barcode","chr12_25245351_G","chr5_175478615_A")

metadata <- tuft.merged@meta.data %>% as.data.frame()
metadata <- left_join(metadata, kras.mut, "Barcode" )

metadata %>% group_by(Patient_ID) %>% count(chr12_25245351_G)
chr12_25245351_G <- metadata$chr12_25245351_G
names(chr12_25245351_G) <- metadata$S_CB
tuft.merged <- AddMetaData(object = tuft.merged, metadata = chr12_25245351_G, col.name = "KRAS.mut")
col <- c("red","red","black","black","grey")
DimPlot(tuft.merged, group.by = "KRAS.mut", cols = col)
```

# Transcriptome profiling of Tuft and ADM-derived cells
```{r}
ADM <- subset(sts.non.malignant, subset = Lv.2.Anno %in% c("Tuft","Acinar","Acinar-Ductal","Ductal epithelial", "Ductal_MUC5B+")) #"Tuft.persistent"))
DimPlot(ADM, group.by = "Lv.2.Anno", label = T)

ADM.col <- c("#6A3D9A","#D76C7C","#5E7D9A","#DCC179" ,"#1B9E77" )
ADM.color <- setNames(object = ADM.col, nm=c("Tuft","Acinar", "Acinar-Ductal",
                                             "Ductal epithelial", "Ductal_MUC5B+"))
```

## Pseudobulk - based on annotated clusters on sts.non.malignant object
```{r}
# Note: not including persistent tuft in this pseudobulk as it comes from only 1 patient
DefaultAssay(ADM) <- "RNA"
adm <- ADM

# Converting cluster name into numerical
Idents(adm) <- "Lv.2.Anno"
adm@meta.data$Lv.2.Anno <- as.factor(adm@meta.data$Lv.2.Anno)
print(levels(adm@meta.data$Lv.2.Anno))
levels(adm@meta.data$Lv.2.Anno) <- 1:length(levels(adm@meta.data$Lv.2.Anno))
print(1:length(levels(adm$Lv.2.Anno)))

ClusterSub <- as.factor(adm@meta.data$Lv.2.Anno) 
PatClust <- paste(adm$Patient_ID, ClusterSub, sep = "_Clst") #labelling each cell with Sample#_Cls#

counts <- adm@assays$RNA@counts #dgCMatrix
counts <- as.matrix(counts)
counts <- t(rowsum(t(counts), group=PatClust))

library(limma)
library(edgeR)
count.matrix <- DGEList(counts) 
count.matrix$samples$Patient <- gsub("_Clst.*$","", colnames(count.matrix)) #adding a new column, named Patient 
count.matrix$samples$Cluster <- as.numeric(gsub("^.*_Clst","",colnames(count.matrix))) #adding a new column, named Cluster
count.matrix$samples$group <- count.matrix$samples$Cluster
count.matrix #count matrix

ncls <- length(table(ClusterSub))
N <- 1:ncls #focus on all clusters
ClstSub <- count.matrix[,count.matrix$samples$Cluster %in% N] 
keep <- filterByExpr(ClstSub, min.count=3, min.total.count=10) #min=10, min.total.count=20
ClstSub <- ClstSub[keep, , keep=F]
ClstSub <- calcNormFactors(ClstSub) # to calculate normalisation factors to a algin columns of a count matrix
ClstSub$samples$group

plotMDS(ClstSub, pch = 16, col = ADM.col[ClstSub$samples$group], main="")
 
#Design matrix
Cls <- as.factor(ClstSub$samples$Cluster)
Pat <- factor(ClstSub$samples$Patient)
design <- model.matrix(~ Cls + Pat)

#Estimate NB dispersion & QL dispersion
ClstSub <- estimateDisp(ClstSub, design = design)
qfit <- glmQLFit(ClstSub, design)

contr <- makeContrasts(
    Cls1 = -(Cls2+Cls3+Cls4+Cls5)/4,
    Cls2 = Cls2-(Cls3+Cls4+Cls5)/4,
    Cls3 = Cls3-(Cls2+Cls4+Cls5)/4,
    Cls4 = Cls4-(Cls2+Cls3+Cls5)/4,
    Cls5 = Cls5-(Cls2+Cls3+Cls4)/4,
    levels=design)

ctest <- list()
for (i in 1:ncls) ctest[[i]] <- glmQLFTest(qfit, contrast = contr[,i])

#log-CPM
prior.count <- 1
ClstSub.cpm <- edgeR::cpm(ClstSub, log=T, prior.count=prior.count) #calculating cpm

# Remove batch effect on log-CPM
logCPM_no_batch <- removeBatchEffect(ClstSub.cpm, batch=Pat, group = Cls) # for visualisation purposes only
plotMDS(logCPM_no_batch, pch = 16, col = ADM.col[ClstSub$samples$group], main="Pseudo-bulk samples")

### Extracting data table ###
filtered_clusters <- list()

# Loop over the clusters
for (i in 1:5) {
  cluster <- ctest[[i]]$table %>% mutate(cluster = i)
  cluster$gene <- rownames(cluster)
  filtered_clusters[[i]] <- dplyr::filter(cluster, abs(logFC) >= 1, PValue < 0.05)
}

adm.cellnames.pseudobulk %>% mutate(cell_names = case_when(
                                         cluster == "1" ~ "Acinar",
                                         cluster == "2" ~ "Acinar-Ductal",
                                         cluster == "3" ~ "Ductal epithelial",
                                         cluster == "4" ~ "Ductal_MUC5B+",
                                         cluster == "5" ~ "Tuft",
                                         )) -> adm.cellnames.pseudobulk

write.csv(adm.cellnames.pseudobulk, file = here(data.path, "Supplmentary_Table_3.csv", row.names = F))
```


# PAGA analysis
## ADM process in ADM samples - PAGA analysis
```{r}
sts.simspec.sct <- readRDS("../sts.simspec_css.sample.sct.Aug24.RDS")
ADM.Pat <- subset(sts.simspec.sct, subset = Patient_ID %in% c("2285","2215","2221","2156") &
                                            simspec_cell_class %in% c("Acinar","PDAC","Ductal_epithelial","Tuft"))

DefaultAssay(ADM.Pat) <- "RNA"
Idents(ADM.Pat) <- "Lv.2.Anno"
ADM.Pat
ADM.Pat <- subset(ADM.Pat, subset = Lv.2.Anno == "unassigned", invert = T)

ADM.Pat[['Lv.2.Anno']] <- lapply(ADM.Pat[['Lv.2.Anno']], as.character)

# Convert seurat to Anndata
SaveH5Seurat(ADM.Pat, filename = "~/projects/PDAC_non_malignant_cell_types/data/PAGA/adm_n4_Nov24.h5Seurat", overwrite = T)
Convert("~/projects/PDAC_non_malignant_cell_types/data/PAGA/adm_n4_Nov24.h5Seurat", dest = "h5ad", overwrite = T)
saveRDS(ADM.Pat, file=here(objects.path, "ADM.Pat.RDS"))
```

### Plotting
```{r}
# Fig2a
DefaultAssay(ADM) <- "RNA"
VlnPlot(ADM, features = c("CPB1","CPA1","REG1A",'EGF',
                          "CFTR",'SLC4A4',
                          "MUC5B",'MUC6', 
                          "TRPM5","AVIL","POU2F3",'BMX'), cols = ADM.color, pt.size = 0, group.by = "Lv.2.Anno")

# Sup Fig2a
features <- list()
features$acinar <- c("CPB1","CTRC","CELA3A","PNLIP","PRSS1","CUZD1","CTRB2","CPA2","SPINK1")
features$REG <- c("REG1B","REG3A","REG3G","REG1A")
features$acinar.ductal <- c("BNIP5","PRKD1","MECOM","OLFM4","FGD6","ABLIM2","IL32","CXCL2")
features$ductal <- c("CFTR","SLC4A4","PKHD1","ONECUT2","BICC1","GLIS3","SCTR")
features$ductal.like2 <- c("MUC6","MUC5B","MUC3A","TFF2","CALN1","SERPINA1","IL33")
features$PDAC <- c("CDKN2A","SMAD4","PIK3CA","KRT7","KRT20","KRT19","KRT5","MUC1","CDH1","SDC1","FXYD3","KRT17","KRT8","EPCAM","TFF1","PPARG","CAPN8","TMC5","ACTN4","LMO7")
features$Tuft <- c("TRPM5","POU2F3","AVIL","BMX","SH2D6","SH2D7")

ADM.Pat$Lv.2.Anno <- factor(ADM.Pat$Lv.2.Anno, levels = c("Acinar", "Acinar-Ductal","Ductal epithelial", "Ductal_MUC5B+","PDAC", "Tuft","Tuft.malignant"))
Idents(ADM.Pat) <- "Lv.2.Anno"

DefaultAssay(ADM.Pat)
DotPlot(ADM.Pat, features = features) + RotatedAxis() + scale_color_gradient2(low="darkblue", mid="white", high="darkred")

# Fig 2g
genes <- c("MDK","IL15","IL19",
           "IL17RB","IL13RA1","IL6R","HCK",
           "PVR", "LGALS3","LGALS4","LGALS9")

Idents(ADM) <- "Lv.2.Anno"
DefaultAssay(ADM) <- "RNA"
DotPlot(ADM, features = genes, group.by = "Lv.2.Anno") + RotatedAxis() + scale_color_gradient2(low="darkblue", mid = "white", high = "darkred")
```
