---
title: "CellChat Analysis"
author: "Mai La"
---

# CELLCHAT - Lv.1.Anno
```{r}
library(CellChat)
library(patchwork)
library(Seurat)
options(stringsAsFactors = FALSE)
options(future.globals.maxSize = 8000 * 1024^2)
# CellChat requires two user inputs: one is the gene expression data of cells, and the other is the user assigned cell labels.
# https://htmlpreview.github.io/?https://github.com/jinworks/CellChat/blob/master/tutorial/CellChat-vignette.html
```

## Create input files 
Ref: https://htmlpreview.github.io/?https://github.com/jinworks/CellChat/blob/master/tutorial/CellChat-vignette.html#infer-the-cell-cell-communication-at-a-signaling-pathway-level
```{r}
sts.simspec.final <- readRDS("~/projects/PDAC_non_malignant_cell_types/output/cellbender/Aug24_objects/sts.simspec.final.RDS")
DefaultAssay(sts.simspec.final) <- "RNA"
Idents(sts.simspec.final) <- "Lv.1.Anno"
sts.simspec.final <- subset(sts.simspec.final, subset = Lv.1.Anno %in% c("Stromal (rare subtypes)", "unassigned","PDA_OLGC"), invert=T)
sts.seurat <- subset(sts.simspec.final, downsample=500) # downsample parameter here means 500 cells per idents
```

## Creating CellChat object
```{r}
cellchat <- createCellChat(object = sts.seurat, group.by = "Lv.1.Anno", assay = "RNA") # require lognormalise data
cellchat@DB <- CellChatDB.human

# Preprocessing the expression data for cell-cell communication
cellchat <- subsetData(cellchat, features = NULL) # This step is necessary even if using the whole database
future::plan("multisession", workers = 20) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- computeCommunProb(cellchat, type = "truncatedMean", trim = 0.15) # produce fewer interactions but still detecting strong interaction, trim=0.15 mean 15% of cell express
```

## TO EXAMINE PVR 
```{r}
pathways.show <- c("PVR")
# CIRCLE PLOT - FIG 2E
par(mfrow=c(1,1))
netVisual_aggregate(cellchat, signaling = pathways.show, layout = "circle")
saveRDS(cellchat, file = "~/projects/PDAC_non_malignant_cell_types/output/cellbender/Aug24_objects/cellchat_Lv.1_rerun.RDS")
```

```{r}
ptm = Sys.time()
cellchat <- computeCommunProbPathway(cellchat)
# Compute the network centrality scores
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups

# PVR SIGNALLING PATHWAY NETWORK - FIG 2F
netAnalysis_signalingRole_network(cellchat, signaling = pathways.show, width = 15, height = 2.5, font.size =6)
```

# CELLCHAT - Lv.2.Anno
```{r}
DefaultAssay(sts.simspec.final) <- "RNA"
Idents(sts.simspec.final) <- "Lv.2.Anno"
sts.simspec.final <- subset(sts.simspec.final, subset = Lv.2.Anno %in% c("unassigned","PDA_OLGC"), invert=T)
sts.seurat <- subset(sts.simspec.final, downsample=500) # downsample parameter here means 500 cells per idents
```

## Creating CellChat object
```{r}
cellchat <- createCellChat(object = sts.seurat, group.by = "Lv.2.Anno", assay = "RNA") # require lognormalise data
cellchat@DB <- CellChatDB.human
showDatabaseCategory(cellchat@DB)

# Preprocessing
cellchat <- subsetData(cellchat, features = NULL) # This step is necessary even if using the whole database
future::plan("multisession", workers = 10) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
saveRDS(cellchat, file = "~/projects/PDAC_non_malignant_cell_types/output/cellbender/Aug24_objects/cellchat_Lv.2_rerun.RDS")
```

## Infer cell-cell communication network
https://htmlpreview.github.io/?https://github.com/jinworks/CellChat/blob/master/tutorial/CellChat-vignette.html#infer-the-cell-cell-communication-at-a-signaling-pathway-level
```{r}
cellchat <- computeCommunProb(cellchat, type = "truncatedMean", trim = 0.15)
cellchat <- computeCommunProbPathway(cellchat)

cellchat <- aggregateNet(cellchat)
groupSize <- as.numeric(table(cellchat@idents))
```

### Tumour vs TAMs_SPP1_OLR1
```{r}
# Fig 3h
df.net <- subsetCommunication(cellchat, sources.use = c("TAMs_SPP1+","TAMs_OLR1+"), targets.use = c("PDAC")) 
df.net <- df.net %>% dplyr::filter(!grepl("COLLAGEN|LAMININ|SEMA3|SEMA4", pathway_name, ignore.case=TRUE))

netVisual_bubble(cellchat, sources.use = c("TAMs_SPP1+","TAMs_OLR1+"), targets.use = "PDAC",
                 signaling = c("TGFb","ACTIVIN","EGF","SPP1","GRN","CADM","APP"),
                 remove.isolate = FALSE)
```

```{r}
### Tumour vs TAMs_SPP1_OLR1
```{r}
# Fig 3h
df.net <- subsetCommunication(cellchat, sources.use = c("Tuft"), targets.use = c("Ductal MUC5B+","PDAC")) 
df.net <- df.net %>% dplyr::filter(!grepl("COLLAGEN|LAMININ|SEMA3|SEMA4", pathway_name, ignore.case=TRUE))

netVisual_bubble(cellchat, sources.use = c("Tuft"), targets.use = c("Ductal MUC5B+","PDAC"),
                 signaling = c("FGF","MK"),
                 remove.isolate = FALSE)
```

```

```{r}
NRT <- subsetCellChat(
  cellchat,
  cells.use = NULL,
  idents.use = c("TAMs_SPP1+","TAMs_OLR1+","NK","Tuft","CD8+ T:ITGAE+","PDAC"),
  group.by = NULL,
  invert = FALSE,
  thresh = 0.05
)

df.net <- subsetCommunication(cellchat, sources.use = c("TAMs_SPP1+","TAMs_OLR1+","NK","Tuft","CD8+ T:ITGAE+"), targets.use = c("PDAC")) 
df.net <- df.net %>% dplyr::filter(!grepl("COLLAGEN|LAMININ|SEMA3|SEMA4", pathway_name, ignore.case=TRUE))

netVisual_bubble(cellchat, sources.use = c("TAMs_SPP1+","TAMs_OLR1+","NK","Tuft"), targets.use = "PDAC",
                 signaling = c("TGFb","ACTIVIN","EGF","SPP1","GRN","CADM","APP","GALECTIN","PARs","MK","CEACAM"),
                 remove.isolate = FALSE)

pathways.show = c("PARs","CD96")
netVisual_individual(NRT, signaling = pathways.show, layout = "circle")
```

## CAFs_NRT_TAMsC1QC_MHCII_Lymphoids
```{r}
# Fig 4g
NRT <- subsetCellChat(
  cellchat,
  cells.use = NULL,
  idents.use = c("CAFs_NRT","TAMs_C1QC+","CD3+ T", "CD4+ T:IL7R+","Schwann","TAMs_HLA+","Migratory DCs","cDC1", "cDC2", "CD8+ T:GZMK+"),
  group.by = NULL,
  invert = FALSE,
  thresh = 0.05
)

df <- subsetCommunication(NRT)
df <- df %>% dplyr::filter(!grepl("COLLAGEN|LAMININ|SEMA3|SEMA4", pathway_name, ignore.case=TRUE))

pathways.show = "CD80"
netVisual_individual(NRT, signaling = pathways.show, layout = "circle") 

pathways.show = "CD86"
netVisual_individual(NRT, signaling = pathways.show, layout = "circle")
```

### TILs
```{r}
# Fig 5f
TIL <- subsetCellChat(
  cellchat,
  cells.use = NULL,
  idents.use = c("Migratory DCs","CD4+ T:IL7R+","CD4+ T:FOXP3+", "CD8+ T:GZMK+", "CD8+ T:ITGAE+", "CD3+ T", "NK", "PDAC","TAMs_C1QC+", "TAMs_HLA+", "cDC1"), 
  group.by = NULL,
  invert = FALSE,
  thresh = 0.05
)

df <- subsetCommunication(NRT)
df <- df %>% dplyr::filter(!grepl("COLLAGEN|LAMININ|SEMA3|SEMA4", pathway_name, ignore.case=TRUE))

netVisual_bubble(TIL, sources.use = c("PDAC"), targets.use = "CD8+ T:ITGAE+" , remove.isolate = FALSE) # Not much except PVR-CD226
pathways.show = c("PVR")
netVisual_individual(TIL, signaling = pathways.show, layout = "circle") 
```

### TLS 
```{r}
# Fig 6g
TLS <- subsetCellChat(
  cellchat,
  cells.use = NULL,
  idents.use = c("B","FRCs","FDCs","CD3+ T", "CD4+ T:IL7R+","Migratory DCs","cDC1", "cDC2", "CD8+ T:GZMK+","CD8+ T:ITGAE+","TAMs_C1QC+","Schwann","Endothelial","CAFs_NRT","LECs"),
  group.by = NULL,
  invert = FALSE,
  thresh = 0.05
)

df <- subsetCommunication(TLS)
df <- df %>% dplyr::filter(!grepl("COLLAGEN|LAMININ|SEMA3|SEMA4", pathway_name, ignore.case=TRUE))

pathways.show = c("CCL","CXCL")
netVisual_individual(TLS, signaling = pathways.show, layout = "circle")
```
