---
title: "Sample_processing_Seurat"
author: "Hue M. La & Fernando J. Rossello"
output: html_document
---

Based on the approach described [here](https://satijalab.org/seurat/articles/integration_introduction.html), example described in this [manuscript](https://www.cell.com/cell/fulltext/S0092-8674(19)30559-8)
Primary processing of STS PDAC samples including IPMN using Seurat, QCed using SCRAN/SCATER (OSCA approach)
All data input is placed in `data/`
Plots and figures are placed in `output/plots/`

# Prepare environment
```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, error = F, message = F, results = "hide", eval = T)
bpp <- BiocParallel::MulticoreParam(workers = 28, RNGseed = 1234)
```

## Libraries
```{r import_libraries, include = TRUE, echo=TRUE, results='hide', message=FALSE}
library("Seurat")
library("tidyverse") 
library("patchwork")
library("BiocSingular")
library("BiocParallel")
library("pheatmap")
library("bluster")
library("dichromat")
library("RColorBrewer")
library("clustree")
library('scuttle')

objects.path <- here("output/objects")
```

# Load QCed SCE object.
```{r load_sce_object}
qc.dbl.fltrd.sts.sce <- readRDS("../qc.dbl.fltrd.sts.sce.RDS")

sample.no <- length(unique(qc.dbl.fltrd.sts.sce$Sample))
sample.colours  <- colorRampPalette(brewer.pal(11, "Spectral"))
```

# Create a Seurat object (QCed with Scater)
```{r seurat_object}
rownames(qc.dbl.fltrd.sts.sce) <-
  uniquifyFeatureNames(rownames(qc.dbl.fltrd.sts.sce),
                       rowData(qc.dbl.fltrd.sts.sce)$Symbol)
# Idea from this [post](https://stackoverflow.com/questions/28244123/find-duplicated-elements-with-dplyr)

rowData(qc.dbl.fltrd.sts.sce) %>% as_tibble (rownames = "rownames") %>% dplyr::filter(duplicated(Symbol) | duplicated(Symbol, fromLast = TRUE))

sts.seurat <-
  CreateSeuratObject(
    counts = counts(qc.dbl.fltrd.sts.sce),
    meta.data = colData(qc.dbl.fltrd.sts.sce) %>% as.data.frame()
  )
rm(qc.dbl.fltrd.sts.sce)
gc()
```

# SCTransform 
Normalisation and scaling using "v2" flavour.
Run as source code (`SCTransform.R`).
```{r sctransform, eval=FALSE}

sts.seurat <- SCTransform(
  sts.seurat,
  vst.flavor = "v2",
  method = "glmGamPoi",
  verbose = TRUE) 
```

# RNA assay
```{r rna_assay}
sts.seurat <- 
  NormalizeData(
    object = sts.seurat,
    normalization.method = "LogNormalize",
    scale.factor = 1e4,
    assay = "RNA"
  )

GetAssayData(sts.seurat, slot = "counts", assay = "RNA") %>% .[1:50, 1:50]
GetAssayData(sts.seurat, slot = "data", assay = "RNA")  %>% .[1:50, 1:50]
```

# Dimensionality reduction
## PCA
```{r pca}
sts.seurat <-
  RunPCA(sts.seurat,
         verbose = TRUE,
         assay = 'SCT',
         reduction.name = "pca",
         reduction.key = "PCA_")

DimPlot(sts.seurat,
        label = TRUE,
        reduction = "pca",
        group.by = "Sample", raster = F) + NoLegend()

ElbowPlot(sts.seurat, ndims = 50)

DimHeatmap(
  object = sts.seurat,
  dims = 1:12,
  cells = 500,
  balanced = TRUE
)
```

## UMAP
```{r umap}
library(Seurat)
sts.seurat <- RunUMAP(
    sts.seurat,
    dims = 1:50,
    verbose = TRUE,
    assay = 'SCT',
    reduction = 'pca',
    reduction.name = 'umap',
    reduction.key = 'UMAP_'
  )

DimPlot(
  sts.seurat,
  label = F,
  group.by = "Sample",
  reduction = "umap",
  raster = FALSE)

```

# Clustering 
```{r clustering}
sts.seurat <-
  FindNeighbors(
    sts.seurat,
    reduction = 'pca',
    dims = 1:50,
    verbose = TRUE,
    assay = 'SCT'
  )

sts.seurat <- FindClusters(sts.seurat, algorithm = 3, resolution = seq(0.1, 1, by = 0.1))

library(clustree)
clustree::clustree(sts.seurat) 

Idents(sts.seurat) <- "SCT_snn_res.0.5"
sts.seurat$seurat_clusters <- NULL
sts.seurat$seurat_clusters <- sts.seurat$SCT_snn_res.0.5 

DimPlot(
  sts.seurat,
  reduction = "umap",
  raster = F,
  group.by = "SCT_snn_res.0.5",
  label = T
) +
DimPlot(sts.seurat, reduction = "umap", group.by = 'Sample', raster =F)
```

## QC/Clustering plots
```{r qc_clustering}
FeaturePlot(sts.seurat, features = "nCount_RNA", slot = 'data', reduction = "umap", raster = F) + ggtitle('UMAP highlighting number of UMIs per cell') +
FeaturePlot(sts.seurat, features = "nFeature_RNA", reduction = "umap", raster = F) + ggtitle('UMAP highlighting number of genes per cell') +
FeaturePlot(sts.seurat, features = "subsets_mito_percent", reduction = "umap", raster = F) + ggtitle('UMAP highlighting the percentage of UMIs\nassociated with mitochondrial genes')
```

## Clustree
```{r}
clustree(sts.seurat, prefix = "SCT_snn_res.")
```

# Manual annotation based on marker gene expression
```{r}
DefaultAssay(sts.seurat) <- "RNA"
marker.sets <- list(Malignant_PDAC = c("KRT19","MUC1","KRT17","ACTN4","LMO7"),
                    Malignant_Acinar = c("CPA1","CPA2","PRSS1","RBPJL","CEL"), 
                    Malignant_PDA_Osteoclast = c("DLX5","OMD","CA2"),
                    IPMN = c("MUC2","FGF13","REG4","BRINP3","SPINK4"),
                    IPMN_intestinal = c("CEMIP","TCIM","NR4A2","DNER"),
                    Acinar_Cells = c("CPB1","CELA3A","CTRC","PNLIP","PDIA2"),
                    Epithelium_Duct_Cells = c("CFTR","SLC4A4","PKHD1","ONECUT2","C6","CRP"),
                    Ductal_like_2 = c("MUC6","MUC5B","MUC3A","CALN1","SERPINA1"),
                    Quiescent_stellates = c("PDGFRB","RGS5","TRPC6","PLXDC1","GJC1"),
                    Fibroblasts = c("COL1A1","CALD1", "COL6A3","COL1A2","FN1"),
                    Fibroblastic_reticular = c("HAS1","PDPN","PLA2G2A"),
                    Tuft_Cells = c("TRPM5","POU2F3","AVIL","BMX","SH2D7"),
                    Islet_Cells = c("INS","GCG","NRG1","KCNMB2","SNTG1"),
                    Endothelial_Cells = c("VWF","CD34","EFNB2","PLVAP","FLT1"),
                    B_Cells = c("FCRL1","MS4A1","PAX5"),
                    Plasma_Cells = c("MZB1","SPAG4","PDK1","IGHG1"),
                    T_Cells = c("CD3D","CD3E","CD96","CD2","CD247"), 
                    Mast_Cells = c("TPSAB1", "CPA3", "MS4A2","HPGDS","KIT"),
                    Macrophages = c("CSF1R","SIGLEC1", "CD163", "MS4A6A"))

DotPlot(sts.seurat, features = marker.sets, group.by = "SCT_snn_res.0.5")  + RotatedAxis() + scale_color_gradient2(low="darkblue", mid="white", high = "darkred") # res 0.6 work very well
```

# Rename clusters
```{r}
# Add Cell_Ids & S_CB
sts.seurat$Cell_Ids <- colnames(sts.seurat)
sts.seurat$S_CB <- paste(sts.seurat$Sample, sts.seurat$Barcode, sep = "_")

clusters <- FetchData(sts.seurat, vars = c("Cell_Ids", "seurat_clusters"))

clusters <-
  clusters %>% dplyr::mutate(
    cell_class = dplyr::case_when(
      seurat_clusters %in% c("5") ~ "T/NK",
      seurat_clusters == "32" ~ "Mast",
      seurat_clusters == "1" ~ "Macrophages",
      seurat_clusters == "28" ~ "Plasma",
      seurat_clusters == "26" ~ "B",
      seurat_clusters == "21" ~ "Endothelial",
      seurat_clusters %in% c("22","15") ~ "Endocrine",
      seurat_clusters == "31" ~ "Tuft",
      seurat_clusters == "33" ~ "FRCs",
      seurat_clusters == "27" ~ "Pericytes",
      seurat_clusters == "2" ~ "CAFs",
      seurat_clusters == "30" ~ "Ductal_MUC5B+",
      seurat_clusters == "6" ~ "Ductal_epithelial",
      seurat_clusters == "24" ~ "Acinar",
      seurat_clusters %in% c("13","29") ~ "PDA_OLGC", 
      seurat_clusters == "7" ~ "Intestinal IPMN",
      seurat_clusters %in% c("0","10") ~ "Malignant acinar",
      seurat_clusters %in% c("34","30","25","23","20","19","18","17","16","14","12","11","9","8","4","3") ~ "PDAC")
  )
stopifnot(identical(colnames(sts.seurat),
                    rownames(clusters)))


sts.seurat <- AddMetaData(sts.seurat,
                          metadata = clusters[, "cell_class"],
                          col.name = "cell_class")

DimPlot(sts.seurat,
        group.by = "seurat_clusters", raster = F, label = T) +
  DimPlot(sts.seurat, raster = F, label = T, group.by = "cell_class", repel = T)

# Assign Patient_ID
replicate.id <- FetchData(sts.seurat, vars = c("Sample", "Cell_Ids"))
replicate.id <- dplyr::mutate(replicate.id, Patient_ID = gsub("\\-.*", "", Sample))
patient.id <- replicate.id$Patient_ID
names(patient.id) <- replicate.id$Cell_Ids

stopifnot(identical(names(patient.id), colnames(sts.seurat)))

sts.seurat <- AddMetaData(sts.seurat, metadata = patient.id, col.name = "Patient_ID")
saveRDS(sts.seurat, file = here(objects.path,"Seurat/sts.seurat.RDS"))
```
